### 2022-05-05 providence/draft.tex

% This is a draft of the coming paper "Providence: A Flexible Round-by-Round Risk-Limiting Audit"
% 2021.
%
\documentclass[runningheads]{llncs}
%
\usepackage{float}
\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{./imgs/}}
\usepackage{comment}
\usepackage{listings,hyperref}
\usepackage{cite}

% \usepackage{graphicx}  --> Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

% nice looking audit titles
\newcommand{\Minerva}{\textsc{Minerva}\xspace}
\newcommand{\Providence}{\textsc{Providence}\xspace}
\newcommand{\B}{{{B2}}\xspace}
\newcommand{\R}{{{R2}}\xspace}
\newcommand{\BRAVO}{\textsc{Bravo}\xspace}

\usepackage{color}
\definecolor{green}{rgb}{0.31, 0.78, 0.47}
\newcommand{\fpo}[1]{\textcolor{green}{#1}}

\def\code#1{\texttt{#1}}

\begin{document}
\lstset{language=Python}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Oliver Broadrick\inst{1} \and
%Sarah Morin\inst{1} \and
%Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
%Filip Zag{\'o}rski\inst{3}\inst{4}}
%
%\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University\thanks{Supported in part by NSF Award 2015253} 
\email{odbroadrick@gmail.com}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}
\\
\and Wroclaw University of Science and Technology\thanks{Author was partially supported by Polish National Science Centre contract number DEC-2013/09/D/ST6/03927} and 
\and Votifica
\\
\email{filip.zagorski@gmail.com}
}
%

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
A Risk-Limiting Audit (RLA) draws paper election ballots in subsequent samples called rounds. 
After each round, a rigorous error criterion decides whether to stop (confirming the announced outcome) or continue (draw another round).
The most commonly used RLA, \BRAVO (Stark), requires the fewest ballots when round sizes are 1. 
Recent RLA \Minerva requires fewer ballots than \BRAVO but requires round sizes to be fixed before the audit begins.
We present \Providence, and audit that has similar efficiency to \Minerva and supports choosing round sizes during the audit.
We prove \Providence is Risk-Limiting and resistant to an adversary who can choose subsequent round sizes given previous samples.
We present simulations in which \Providence has similar efficiency to \Minerva.
Finally, we describe the pilot use of \Providence by the Rhode Island Board of Elections.
Our implementation of \Providence in the R2B2 software library provides standard functionality 
plus a function which gives the probability with which an announced loser will appear to be winning
after a round, a potentially useful metric to election officials when choosing round sizes.
% and integrated into Arlo, election audit software used across the US.
\keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
\end{abstract}
%
%
\section{Introduction}
\label{sec:intro}
\input{sections/intro}

\section{Related work}
\label{sec:related}
\input{sections/related}

\section{\Providence}
\label{sec:prov}
\input{sections/prov}

\section{Simulations}
\label{sec:sims}
\input{sections/sims}

\section{Pilot}
\label{sec:pilot}
\input{sections/pilot}

\section{Conclusion}
\label{sec:conc}
\input{sections/conc}

%\section{Acknowledgements}
%\label{sec:ack}
%This research was supported in part by NSF Awards 2015253
%and 1421373.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{audits}
%
\end{document}