### 2022-07-20 providence/sections/conc.tex
more

%Conclusions
A rigorous tabulation audit is an important part of a secure election. Ballot polling RLAs are commonly used and simple, not relying on special election equipment like comparison RLAs. We present \Providence which is the most efficient and secure ballot polling RLA, as efficient as \Minerva  and flexible as \BRAVO. We present proofs and simulation results to verify the claimed properties of \Providence, and we provide an open source implementation of the stopping condition and useful related functionality.