### 2022-10-11 providence/sections/workload.tex
more

Some election audits have benefited from a one-and-done approach: draw a large sample with high probability of stopping in the first round and usually avoid a second round altogether. This is appealing for two reasons. Firstly, rounds have some overhead in both time and effort. Thus the time and personhours of an audit grows not just with the number of ballots sampled but also with the number of rounds. Secondly, smaller first round sizes give a higher probability that the result after the first round is misleading: the true winner has fewer votes in the audit sample than some other candidate. On the other hand, a one-and-done audit may draw more ballots than are necessary; a more efficient round schedule could require less effort and time pre-certification. To evaluate the quality of various round schedules, we construct a simple workload model. Under this model we show how optimal round schedules can be chosen. We provide software that can be used by election officials to choose round schedules based on estimates of the model parameters like maximum allowed probability of a misleading audit sample.

As an example, we consider the US Presidential contest in the 2016 Virginia statewide general election. This contest had a margin of $5.3\%$ between the two candidates with the most votes.
Analytical approximation of the expected audit behavior (quantities like expected total number of ballots sampled or total number of rounds) is challenging because the number of possible sequences of samples grows exponentially with the number of rounds. 
%A very rough approximation scheme is possible and may be useful when choosing round sizes in practice. We implement such a scheme, available at \cite{software}.
%We will use the more standard approach of simulations to give an example here.
Therefore we use the typical approach of simulations, again with risk limit $10\%$.
We simulate audits considering each candidate with a column in the results available at the Virginia Department of Elections website including irrelevant ballots.
We consider a simple round schedule, in which each round is selected to give the same probability of stopping, $p$. That is, if the audit does not stop in the first round, we select a second round size which, given the sample drawn in the first round, will again have a probability of stopping $p$ in the second round. Note that since there are multiple candidates, we compute the minimum round size to achieve stopping probability $p$ for each sub-audit between the winner and one of the losers, and we then select the largest such minimum round size and scale it up according to the proportion of the total ballots that are relevant to that sub-audit. For this round schedule scheme, a one-and-done audit is achieved by choosing large $p$, say $p=.9$ or $p=.95$.\footnote{For this particular round schedule scheme, computing the expected number of rounds is straightforward analytically, but the expected number of ballots is still difficult, and so we use simulations.} We run $10^4$ trial audits for each value of $p$, assuming the announced results are correct.

\subsection{Person-hours}

\noindent\textbf{Average total ballots.} 
The simplest workload models are a function of just the total number of a ballots sampled. Figure~\ref{fig:avg_bals} shows the average total number of ballots sampled in our simulations for each value of $p$, which gives an estimate of the expected total number of ballots.
Figure~\ref{fig:avg_bals_ratio} gives the same number as a ratio of the \Providence values.
It is straightforward to show that \Providence and both forms of \BRAVO collapse to the same test in the case where each round is a single ballot. Figures~\ref{fig:avg_bals} and \ref{fig:avg_bals_ratio} show that for larger stopping probabilities $p$ (i.e. larger rounds), \Providence requires fewer ballots on average. In particular, the savings of \Providence become larger as $p$ increases; for $p=.95$, EoR \BRAVO and SO \BRAVO require more than $2$ and $1.4$ times as many ballots as \Providence respectively.

\begin{figure}
\includegraphics[width=.5\textwidth]{avg_bals.png}
\caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
\label{fig:avg_bals}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{avg_bals_ratio.png}
\caption{The total number of ballots sampled on average in our simulations given as a fraction of those sampled by \Providence, for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
\label{fig:avg_bals_ratio}
\end{figure}

\noindent\textbf{Round overhead.} 
It is clear that average number of ballots alone is an inadequate workload measure. 
(Consider a state conducting its audit by selecting a single ballot at random, 
notifying just the county where the ballot is located, and then waiting to hear back for the manual interpretation of the ballot before moving on to the next one. 
This of course is inefficient and is why audits are actually performed in rounds.)

In a US state-wide RLA, the state organizes the audit by determining the random sample and communicating with the counties, but election officials at the county level physically sample and inspect the ballots from their precincts. 
Therefore each audit round requires some number of person-hours for set up and communication between state and county. This overhead for a round includes choosing the round size, generating the random sample, and communicating that random sample to the counties, as well as the communication of the results back to the state afterwards.

Consequently, we now consider a model with a constant per-ballot workload $w_b$ and a constant per-round workload $w_r$.
So for an audit with expected number of ballots $E_{b}$ and expected number of rounds $E_{r}$, we estimate that the workload $W$ of the audit is
\begin{equation}
W(E_b,E_r) = E_b w_b + E_r w_r + C
\label{eq:round_workload}
\end{equation}

Note there is also some constant overhead of workload for the whole audit, namely $C$ in Equation~\ref{eq:round_workload}, which we take to be zero in our examples but could be used by election officials for their estimates.
For simplicity, (and without loss of generality), we assume the per ballot workload is one, $w_b=1$. Then the per round workload $w_r$ tells us the workload of a round as a number of ballots. We begin with $w_r=1000$ as a conservative example. 
That is, we set the overhead of a round equal to the workload of sampling $1000$ ballots. Based on available data\cite{RI-report}, the time retrieving and analyzing each individual ballot is on the order of $75$ seconds which means that $w_r=1000$ is equivalent to roughly $20$ person-hours of workload. This corresponds to about $15$ minutes being spent per round in each of the $133$ counties of Virginia, a clearly conservative workload estimate. 
As shown in Figure~\ref{fig:with_round_workload}, lower average workloads are achieved by selecting higher stopping probability; \Providence achieves the lowest minimum average workload is achieved at roughly $0.7$.

\begin{figure}
\includegraphics[width=.5\textwidth]{with_round_workload.png}
\caption{For workload parameters $w_b=1$ and $w_r=1000$, this plot shows the expected workload for various round schedule parameters $p$. Expected workload is found using Equation~\ref{eq:workload} and the average number of ballots and rounds in our simulations as the expected number of ballots and rounds.}
\label{fig:with_round_workload}
\end{figure}

Importantly, this gives us a way to estimate the expected workload, as well as which round schedule value $p$ achieves it, for arbitrary round workload. For each round workload $w_r$, we produce a dataset analagous to that of Figure~\ref{fig:with_round_workload} and then find the minimum average workload achieved for each of the audits and its corresponding stopping probability $p$. 
Figure~\ref{fig:optimal_workloads} shows the optimal achievable workload for a wide range of per round workloads. For very low round workloads, the workload function approaches just the total number of ballots, and so, as seen in Figure~\ref{fig:abv_bals}, the three approaches differ by less. On the other hand, for extremely larger values of round workload, the average number of ballots has little impact on the workload function, and so the three audits again have similar values. For more reasonable values of the round workload $w_r$, SO \BRAVO and EoR \BRAVO achieve minimum workload roughly $1.1$ and $1.3$ times greater than that of \Providence.
Figure~\ref{fig:optimal_ps} shows the corresponding round schedule parameters $p$ that achieve these minimal workloads. As expected, a overhead for each round means that larger round sizes are needed to achieve an optimal audit, and so for all three audit $p$ increases as a function of $w_r$. Notice that \Providence is generally above and to the left of SO \BRAVO, and SO \BRAVO is generally above and to the left of EoR \BRAVO. This relationship reflects the fact that for the same round workload, \Providence can get away with a larger stopping probability because it requires fewer ballots.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_workloads.png}
\caption{For varying round workload $w_r$, the optimal average workload achievable by each audit, as a fraction of the \Providence values. (We show the value for varying $w_r$ since the value of $w_r$ may differ significantly from case to case.) Note that values lower than $10^3$ are probably unrealistic in US statewide contests; in Virginia, $w_r=10^3$ corresponds to only about $15$ minutes of persontime per county as the overhead per round.}
\label{fig:optimal_workloads}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_ps.png}
\caption{The optimal (workload-minimizing) stopping probability $p$ for varying workload model parameters $w_r$. With $w_b=1$, the varying value of $w_r$ can equivalently be thought of as the ratio of the workload of a round to the workload of a ballot. (Note that the steps in this function are a consequence of our subsampling the workload function. That is, the workload-minimizing value of $p$ for each $w_r$ is only allowed to take on values at increments of $0.5$.)}
\label{fig:optimal_ps}
\end{figure}

\noindent\textbf{Precinct overhead.} For a more complete model, we can also introduce container-level workload. The time to sample a ballot from an entirely new box is typically greater than to sample a ballot from an already-open box. Based on a Rhode Island pilot RLA report\cite{RI-report}, this may mean that a ballot from a new container requires roughly twice the time as a ballot from an already-opened container. Typically available election results give per-precinct granularity of vote tallies, rather than individual container information. In Virginia, however, most precincts have a single ballot scanner whose one box has sufficient capacity for all the ballots cast in that precinct anyways, and so we model the per-container workload with an additional per-precinct constant workload, $w_p$. In this model, the workload estimate incurs an additional workload of $w_p$ every time a precinct is sampled from for the first time in a round. That is, let $E_{pi}$ be the expected number of distinct precincts sampled from in round $i$, and let $E_p=\sum_i E_{pi}$. Then the new model is
\begin{equation}
W(E_b, E_r, E_p) = E_b w_b + E_r w_r + E_p w_p + C
\label{eq:round_and_precinct_workload}
\end{equation}

We can again explore the minimum achievable workloads under this model, as shown in Figure~\ref{fig:optimal_workload_precinct_workload_ratio}.

\begin{figure}
\includegraphics[width=.5\textwidth]{optimal_workload_precinct_workload_ratio.png}
\caption{The optimal average workload of the audits we simulated using the workload Equation~\ref{eq:round_and_precinct_workload} for varying $w_p$, given as a fraction of the value for \Providence. Similar to Figure~\ref{optimal_workloads}, we show a generous range of values for the workload variable, $c_p$ in this case. If the time for a single ballot is $75$ seconds, then $c_p=50$ corresponds to over an hour of extra time to sample a ballot from a new container.}
\label{fig:optimal_workload_precinct_workload_ratio}
\end{figure}

\subsection{Real time}
Given tight certification deadlines\footnote{Virginia recently passed legislation requiring pre-certification RLAs TODO check exactly.}, the total real time to conduct the RLA is also an important factor to consider when planning audits.
Because each county can sample ballots for the same round concurrently, the total real time for a round depends only on the slowest county. 
In Virginia, Fairfax County typically has the most votes cast by a significant difference; in the contest we consider, Fairfax County had ~551 thousand votes cast, more than double the ~203 thousand of second-highest Virginia Beach City.
Consequently, we model the expected total real time $T$ of an audit using just the largest county, and we define analagous variables for the expected values in just the largest county.
For the largest county, let the expected total ballots sampled be $\bar E_b$, the expected number of rounds $\bar E_r$, and the expected number of distinct precinct samples summed over all rounds be $\bar E_c$.
Similarly, we use real time per-ballot, per-round, and per-precinct workload variables, $t_b$, $t_r$, and $t_p$. So the real time of the audit is estimated by
\begin{equation}
T(\bar E_b, \bar E_r, \bar E_p ) = \bar E_b t_b + \bar E_r t_r + \bar E_p t_p
\label{eq:real_time}
\end{equation}

As before, we can use our simulations to estimate $\bar E_b$, $\bar E_r$, and $\bar E_p$ using the corresponding averages over the trials. 
Available data to estimate values for $t_b$, $t_r$, and $t_p$ is limited, and so we take as an example the values $t_b=75$ seconds, $t_r=3$ hours, and $t_p=75$ seconds.\footnote{The value $t_b=75$ seconds corresponds to a serial retrieval and interpretation of the ballots based on the \cite{RI-report} timing, $t_p=75$ seconds corresponds to the approximate doubling in time for new-box ballots as reported in \cite{RI-report} in the ballot-level comparison timing data, and $t_r=3$ hours is just a guess at an approximate order for this variable.} In practice, election officials could use our software and their own estimates of these values to explore choices for round schedules. Figure~\ref{fig:real_time} shows how the esimated real time for these values differs as a function of $p$. It should be noted that real values of $t_b$, $t_r$, and $t_p$ will vary greatly based on the number of parallel teams retreiving and checking ballots, the distribution of ballots and containers both in number and physical space, and other factors. We provide Figure~\ref{fig:real_time} only as an example of the general shape and behavior of this function. Use of this optimal scheduling tool would depend on parameter estimates tailored to each case.

\begin{figure}
\includegraphics[width=.5\textwidth]{real_time.png}
\caption{The real time as estimated by Equation~\ref{eq:real_time} for varying $p$ with expected values as estimated by our simulations.}
\label{fig:real_time}
\end{figure}

\subsection{Misleading samples}

Unfortunately, efficiency alone is not sufficient for planning audits. In the US today, election officials need to make legitimate safety considerations.
When drawing a random sample of the ballots, it is always possible that the tally of the sample provides misleading information. In a random sample, a true loser may recieve more votes than the true winner. This happens more often when the sample sizes are small.
In these RLAs, a misleading sample in an early round is dealt with by drawing more ballots (moving on to another round), but in practice the implications of this approach may be dangerous.

Imagine that Alice beats Bob in an election contest both truly and by the announced results, but Bob's supporters are insistent he really won. When election officials carry out the RLA, they choose a small first round size in the hopes of achieving an efficient audit by getting to stop sooner. After the first round, by chance, there are more votes for Bob than for Alice in the sample. Bob's supporters celebrate their victory that the audit has in fact revealed that Bob really won, but the election officials have to explain that they are moving on to a second round. After the second round, there are more votes in the sample for Alice and sufficiently many that the risk limit is met and the audit now ends confirming the announced result that Alice won. This is an undesirable situation.

We introduce the notion of a \emph{misleading sample}, any cumulative sample which, assuming the announced outcome is correct, contains more ballots for a loser than for the winner.
We can again use our simulations to gain insight into the frequency of \emph{misleading samples}.
For each stopping probability $p$, Figure~\ref{fig:misleading} gives the proportion of simulated audits that had a \emph{misleading sample} at any point. 
Notably, this proportion is as high as 1 in 5 for the smaller stopping probability round schedules.
Accordingly, we introduce a new parameter to our audit-planning tool, the maximum acceptable probability that the audit is misleading, the \emph{misleading limit}.

In Figure~\ref{fig:misleading}, horizontal lines are included to show \emph{misleading limits} of $0.1$, $0.01$, and $0.001$.
To achieve a probability of a misleading sample of at most $0.1$, a round schedule with at least roughly $p=.3$ is needed.
To achieve a probability of misleading of roughly $0.01$, a round schedule with $p=0.8$ is neeeded, and to achieve a probability of misleading of roughly $0.001$, a round schedule with $p=0.95$ is needed.
It is not unreasonable to think that election officials might choose a \emph{misleading limit} of $0.01$, or smaller, given the state of public perception of election security in the US and the associated threats of violence.
Consequently, the desired \emph{misleading limit} may be a deciding constraint in the choice of round schedule.

\begin{figure}
\includegraphics[width=.5\textwidth]{misleading_limits.png}
\caption{The proportion of simulated \Providence audits that had a \emph{misleading sample} in any round.}
\label{fig:misleading}
\end{figure}

If election officials wish to enforce a \emph{misleading limit} for all the rounds, our simulation analysis could help. On the other hand, for a given round, it is straightforward to compute analytically the probability that a loser has more votes than the winner in the sample. Table~\ref{tab:misleading} shows for various margins the minimum first round size $n$ that guarantees a probability of a \emph{misleading sample} at most $M\in\{0.1,0.01,0.001\}$. For all values of $M$ and all margins, \Providence achieves a higher probability of stopping than either EoR \BRAVO or SO \BRAVO. 
    As seen in the Table~\ref{tab:misleading}, to enforce $M=0.01$ requires minimum round sizes with at least roughly a $0.8$ probability of stopping in the first round. Even if the most efficient audit schedule (by either workload or real time measures) would use a lower stopping probability $p$ to choose the first round size, the election officials may opt to use this constraint on the probability of a \emph{misleading sample} as the deciding factor in planning their audits.

\begin{table}
\center
\begin{tabular}{ |l|l|r|c|c|c| }
\hline
$M$ & margin & $n$ & Prov & SO & EoR \\
\hline
0.1&0.25&25&0.221&0.152&0.115\\
&0.2&41&0.178&0.169&0.105\\
&0.15&73&0.202&0.186&0.141\\
&0.1&163&0.222&0.182&0.107\\
&0.05&657&0.227&0.192&0.127\\
&0.04&1027&0.237&0.193&0.124\\
&0.03&1825&0.246&0.194&0.124\\
&0.02&4105&0.246&0.195&0.124\\
&0.01&16423&0.246&0.196&0.124\\
\hline
0.01&0.25&85&0.792&0.707&0.559\\
&0.2&133&0.826&0.71&0.593\\
&0.15&239&0.817&0.712&0.549\\
&0.1&539&0.805&0.717&0.567\\
&0.05&2163&0.817&0.721&0.569\\
&0.04&3381&0.82&0.722&0.563\\
&0.03&6011&0.824&0.723&0.573\\
&0.02&13527&0.824&0.723&0.57\\
&0.01&54117&0.824&0.724&0.57\\
\hline
0.001&0.25&149&0.962&0.889&0.783\\
&0.2&235&0.963&0.89&0.768\\
&0.15&421&0.958&0.894&0.801\\
&0.1&951&0.958&0.894&0.793\\
&0.05&3815&0.96&0.896&0.785\\
&0.04&5965&0.961&0.896&0.791\\
&0.03&10607&0.961&0.897&0.787\\
&0.02&23869&0.962&0.897&0.787\\
&0.01&95491&0.962&0.897&0.787\\
\hline
\end{tabular}
\caption{For various margins, this table gives the minimum first round size $n$ to achieve at most a probability $M$ of a \emph{misleading sample} in the first round. The corresponding stopping probabilities of \Providence, SO \BRAVO, and EoR \BRAVO are given for each value of $n$.}
\label{tab:misleading}
\end{table}

\textbf{Misleading SO \BRAVO sequences.} As we consider the idea of misleading samples, it is noteworthy that SO \BRAVO suffers from a different and unique type of misleading result.

After drawing a cumulative $n>1$ ballots in a round, some number $k$ of them are votes for the announced winner. There are $\binom{n}{k}$ possible sequences of ballots which can lead to such a sample. Given a value of $k$, however, the particular sequence of the sample that led to that value of $k$ contains no additional information about whether the sample is more likely under the alternative or null hypotheses. That is to say, $\Pr[K=k|H_a]$ and $\Pr[K=k|H_0]$ have the same value regardless of the sequence.
Despite this, the SO \BRAVO RLA stopping condition is not just a function of $n$ and $k$ but also a function of the sequence, the selection order. In particular, if the sequence of ballots is such that the standard \BRAVO stopping condition was met for some $n'<n$ and corresponding $k'<n$, the audit will stop, even if by the end of the sequence the values $k$ and $n$ no longer meet the \BRAVO condition. We refer to such sequences which stop under SO \BRAVO, but not under EoR \BRAVO, as \emph{misleading sequences}. To be clear, this is not a mathematical issue; stopping in such cases is still a correct application of Wald's SPRT result\cite{wald}. The misleading nature of such stoppages is the note we are making. This is another case where election officials might have difficulty explaining the misleading situation to the public.

As an example, consider

Now recall the pilot \Providence RLA performed in Providence, Rhode Island described in Section~\ref{sec:pilot}. The sample drawn...

It is easy to use our simulations to see how often this occurs by checking whether the final cumulative sample of each SO \BRAVO trial meets the EoR \BRAVO stopping condition and counting those which do not. Figure~\ref{so_misleading} shows the proportion of simulated SO \BRAVO audits that stopped with a \emph{misleading sequence}. Unlike the more general \emph{misleading sample} discussed so far, these \emph{misleading sequences} are unique to SO \BRAVO audits, and Figure~\ref{so_misleading} only shows the proportion of audits that stopped with a \emph{misleading sequence}; additional SO \BRAVO audits also contained \emph{misleading samples}.

\begin{figure}
\includegraphics[width=.5\textwidth]{so_misleading.png}
\caption{The total number of ballots sampled on average in our simulations for various round schedules parameterized by $p$ the conditional stopping probability used to select each round size.}
\label{fig:avg_bals}
\end{figure}