### 2022-10-12 providence/sections/adversary.tex
moved adversary example

% \subsection{Motivation for considering adaptive adversary}
% \fpo{(old notes)}

\Minerva was proven to be a risk-limiting audit for a predetermined round schedule.
As explained earlier, it is not clear that \Minerva is risk-limiting if an adversary can 
adaptively select the round schedule as the audit proceeds. In this section we prove that \Providence does not have this problem, and is risk-limiting even when the adversary can choose next round sizes based on knowledge of the current sample.

\begin{definition}[Strategy-Proof RLA]
 An audit $\mathcal{A}$ is a Strategy-Proof Risk Limiting Audit with risk limit $\alpha$ iff for
 all strategies of selecting round schedule and for sample $X$ 
 \[
  \Pr\left[\mathcal{A}(X) = \text{Correct} | H_0\right] \leq \alpha.
 \]

\end{definition}

\begin{lemma}
\Providence is a Strategy-Proof RLA. 
\end{lemma}
\begin{proof}
This property follows from Theorem~\ref{thm:minerva2_is_rla_new} and Lemma~\ref{lemma:markov}. Note that, as described in section \ref{sec:proof}, the proof of the risk-limiting nature of the audit does not rely on round sizes $n_j$ being identical for all values of $k_{j-1}$. 
\end{proof}

% To see the practical implication of this result, we consider a toy example: an RLA of a two-candidate contest with margin $10\%$ and risk limit $10\%$. 
% Suppose a first round of $100$ ballots is drawn and $60$ ballots in the sample are votes for the announced winner. Since $\tau_1(k,p_a,p_0,\ldots$...

% \newpage

% \subsection{Round extensions}

% The end of a round of size $n_1$, with $k^{p_a, p_0, \alpha}_{min, 1}$ and observed ballots for
% the winner $k$.

% The size of the next round is selected to be $n_{r+1}$.
% The \Providence k-mins are computed as follows. 
% 
% (1) The probability distribution at the end of $(r+1)^{st}$ round is computed:
% % \[
% %  s_{i, d, [n_1, \ldots, n_{r+1}]} = \sum_{j = \max\{0, i - (n_{r+1} - n_r)\}}^{\min\{i, k^{p_a, p_0, \alpha}_{min, r}-1\}} Bin(n_{r+1} - n_r, p_d, i - j) \cdot s_{j, d, [n_1, \ldots, n_r]},
% % \]
% \[
%  s_{i, d, [n_1, \ldots, n_{r+1}]} = \sum_{j = \max\{0, i - (n_{r+1} - n_r)\}}^{\min\{i, k^{p_a, p_0, \alpha}_{min, r}-1\}} Bin(n_{r+1} - n_r, p_d, i - j) \cdot s_{j, d, [n_1, \ldots, n_r]},
% \]
% where $p_d = p_0 = \frac{1}{2}$ or $p_d = p_A$.
% 
% (2) The $k^{p_a, p_0, \alpha}_{min, r}$ is found as:
% \[
%  k^{p_a, p_0, \alpha}_{min, r} =  k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} = \min\left\{ k: \sum_{i = k}^{n_r} s_{i, 0, [n_1, \ldots, n_r]} \le \alpha \sum_{i = k}^{n_r} s_{i, A, [n_1, \ldots, n_r]} \right\}.
% \]

% 
% \begin{definition} 
% Let $[n_1, \ldots, n_r]$ be the round schedule of an audit that has not stopped by the round $r-1$. Let us define 
% \begin{small}
% \begin{equation}\label{eq:kMin}
% k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]} =
%   \min\left\{k : \omega_r(k, k_{r-1},p_a,p_0,n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\}.
% %  \min\left\{k : \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1}) \geq \frac{1}{\alpha}  \right\}$
% \end{equation}
% \end{small}
% Then if $k_r \geq k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]}$ then the result of the audit is Correct (\textit{i.e.,} stopping condition in Definition~\ref{def:minervatwo} holds).
% \end{definition}
% 
% 
% 
% \begin{lemma}
% Let $[n_1, \ldots, n_{r-1}, n_r]$ be a round schedule for an execution of  \Providence audit that has not stopped
% in any of its first $r-1$ rounds (\textit{i.e.,} for every $j = 1, \ldots, r-1$:
% $k_j < k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_j]}$), then: 
% 
% \[ 
% k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_1, \ldots, n_{r-1}, n_r]} = k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_{r-1}, n_r]}.
% \]
% \end{lemma}
% \fpo{this can be used to prove that \Providence is more efficient than \Minerva and \BRAVO}
% \begin{proof}
% Let $k_{r-1}$ denote the number of ballots drawn for the declared winner up to the round $r-1$ (out of $n_{r-1}$ sampled ballots). The stopping decision for the round $r$ is made as follows:
% 
% \[
%  k^{p_a, p_0, \alpha, k_{r-1}}_{min, r, [n_1, \ldots, n_r]} = \min\left\{k : \omega_{j}(k, k_{r-1}, p_a, p_0, n_r, n_{r-1}) \geq \frac{1}{\alpha}  \right\} = 
% \]
% \[
%   =  k^{p_a, p_0, \alpha, k_{r-1}}_{min, 2, [n_{r-1}, n_r]}
% \]
% 
% \end{proof}
% Then, from (\ref{eq:prov}) one can write it as:
% 
% % \[
% %     \begin{aligned}
% %     \omega_{j}(k, k_{r-1}, p_a, p_0, n_{r}, n_{r-1})
% %     \triangleq\\
% %     \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1})
% %     \end{aligned}
% % \]
% \begin{small}
% \[
% \begin{aligned}
%  k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} =\\ 
%  \min\left\{k : \sigma(k_{r-1},p_a,p_0,n_{r-1})\cdot \tau_1(k-k_{r-1},p_a,p_0,n_r-n_{r-1}) \geq \frac{1}{\alpha}  \right\} =\\
%  \min\left\{k :
% \frac{s_{k_{r-1}, a, n_{r-1}}}{s_{k_{r-1}, 0, n_{r-1}}} \frac{
% \sum_{i = k}^{n_r} s_{i - k_{r-1}, a, [n_r - n_{r-1}]}
% }{
% \sum_{i = k}^{n_r} s_{i - k_{r-1}, 0, [n_r - n_{r-1}]}
% }\geq \frac{1}{\alpha}  \right\}
%  \end{aligned}
%  \]
% \end{small}
% 
% \fpo{should be added: for $k > k_{r-1}$ but it seems to be obvious (from monotonicity of k-min)}
% % \[
% % \sigma(k_{r-1},p_a,p_0,n_{r-1}) = \frac{s_{k_{r-1}, a, n_{r-1}}}{s_{k_{r-1}, 0, n_{r-1}}}
% % \]
% 
% % \begin{small}
% % \begin{equation}\label{eq:kMinLong}
% % k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} =\\ 
% %  \min\left\{k :
% % \frac{s_{k_{r-1}, a, n_{r-1}}}{s_{k_{r-1}, 0, n_{r-1}}} \frac{
% % \sum_{i = k}^{n_r} s_{i - k_{r-1}, a, [n_r - n_{r-1}]}
% % }{
% % \sum_{i = k}^{n_r} s_{i - k_{r-1}, 0, [n_r - n_{r-1}]}
% % }\geq \frac{1}{\alpha}  \right\}
% % \end{equation}
% % \end{small}
% %  = \min\left\{ k: \sum_{i = k}^{n_r} s_{i, 0, [n_1, \ldots, n_r]} \le \alpha \sum_{i = k}^{n_r} s_{i, A, [n_1, \ldots, n_r]} \right\}.
% % \]
% 
% Now, one can rewrite the enumerator and the denominator as  follows ($d \in \{a, 0\}$):
% 
% \begin{small}
% \[
% \begin{aligned}
%  s_{k_{r-1}, d, n_{r-1}} \sum_{i = k}^{n_r} s_{i - k_{r-1}, d, [n_r - n_{r-1}]} =\\
%  \mathsf{Bin}(n_{r-1}, p_d, k_{r-1}) \sum_{i = k}^{n_r} 
%  \mathsf{Bin}({n_r - n_{r-1}, d, i - k_{r-1}}) =\\
%  {n_{r-1} \choose k_{r-1}} p_d^{k_{r-1}} p_d^{n_{r-1} - k_{r-1}}
%  \sum_{i = k}^{n_r} 
%  {n_r - n_{r-1} \choose i - k_{r-1}} p_d^{i - k_{r-1}} p_d^{n_r - n_{r-1} - (i - k_{r-1})} =\\
%   \sum_{i = k}^{n_r} {n_{r-1} \choose k_{r-1}}
%  {n_r - n_{r-1} \choose i - k_{r-1}} p_d^{i} p_d^{n_r - i} =
%  \sum_{i = k}^{n_r} {n_{r} \choose i}
%  p_d^{i} p_d^{n_r - i} = \sum_{i = k}^{n_r} s_{i, d, [n_r]}.
%  \end{aligned}
% \]
% \end{small}
% 
% \begin{small}
% \[
% k^{p_a, p_0, \alpha}_{min, r, [n_1, \ldots, n_r]} =\\ 
%  \min\left\{k :
% \frac{
% \sum_{i = k}^{n_r} s_{i, a, [n_r]}
% }{
% \sum_{i = k}^{n_r} s_{i, 0, [n_r]}
% }\geq \frac{1}{\alpha}  \right\}
% = k^{p_a, p_0, \alpha}_{min, 1, [n_r]}
% \]
% \end{small}
% \fpo{There is no need to put $r$ in $k_{min}$ definition $[n_1, \ldots, n_r]$ is enough}
% 
% \end{proof}

% \fpo{Corollary 1: For RO-Bravo you need to pay attention.}
% 
% \fpo{Corollary 2: For \Providence you can (1) be late for an audit (round schedule does not
% matter much); (2) you can be distracted: only the end-of-round matters}

% \begin{lemma}
%  I
% \end{lemma}

% \begin{theorem}
%  \Providence is risk limiting for adversarial choices of round schedules.
% \end{theorem}
% \begin{proof}
% 
% \end{proof}

\subsection{Efficiency}
\begin{lemma}
% For any round schedule %$[n_1, \ldots, n_r]$
For any risk-limit $\alpha \in (0, 1)$, for any margin
and for any round schedule $[n_1, \ldots, n_r]$, 
 \Providence RLA is more efficient than \R \BRAVO.
\end{lemma}
\begin{proof}
 Let $[n_1, \ldots, n_r]$ be a round schedule, and  lets assume that \R \BRAVO audit stopped exactly at
 round $r$, 
 after observing $k_1, \ldots, k_r$ ballots for the announced winner in each round respectively. This means that
 the following conditions hold:
 \begin{enumerate}
  \item for all $j = 1, \ldots, r-1$: 
%   $k_j < k^{p_a, p_0, \alpha}_{min, \BRAVO, n_j}$,
  $\frac{P[{K_j = k_{j}} | H_a, n_j]}{P[K_j = {k_{j}} | H_0, n_j]} < \frac{1}{\alpha}$,
  \item for $j = r$ it holds: % $k_r \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}.$
  $\frac{P[{K_r = k_{r}} | H_a, n_r]}{P[K_r = {k_{r}} | H_0, n_r]} \geq \frac{1}{\alpha}$,
 \end{enumerate}

 From Lemma~\ref{lemma:tau1_increasing} 
 \[
 \tau_1(k, p_a, p_0, n_r) = 
  \frac{P[K_r = {k} | H_a, n_r]}{P[K_r = {k} | H_0, n_r]}
 \]
is a stricly increasing function of $k$ (for $0 < p_0, p_a < 1$ and $n_r > 0$). % for $k \geq k_r$. % (in fact this isk^{p_a, p_0, \alpha}_{min, \BRAVO, n_r} = k^{*}_{n_r}$
% \textit{i.e.,}
% \begin{equation}\label{eq_incr}
% % \frac{1}{\alpha} \leq
%   \frac{P[K_r = {k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} \leq   \frac{P[\mathbf{k_r} + 1 | H_a, n_r]}{P[K_r = {k_r} + 1 | H_0, n_r]} \leq \ldots  \leq \frac{P[\mathbf{n_r} | H_a, n_r]}{P[\mathbf{n_r} | H_0, n_r]}. 
% \end{equation}

If \BRAVO test is satisfied for $k_r$ then:

% From the properties of binomial distribution we obtain:
% \[
%    \frac{P[\mathbf{k_r} | H_a, n_r]}{P[\mathbf{k_r} | H_0, n_r]} = 
%    \frac{P[\mathbf{k_{r-1}} | H_a, n_{r-1}]}{P[\mathbf{k_{r-1}} | H_0, n_{r-1}]}
%    \frac{P[\mathbf{k_r - k_{r-1}} | H_a, n_r - n_{r-1}]}{P[\mathbf{k_r - k_{r-1}} | H_0, n_r - n_{r-1}]} = 
% \]
\[
 \frac{1}{\alpha} \leq \sigma(k_{r}, p_a, p_0, n_{r}) =
\]
\[
 = \sigma(k_{r-1}, p_a, p_0, n_{r-1}) \cdot \sigma(k_r - k_{r-1}, p_a, p_0, n_r - n_{r-1})  \leq
\]
\[
 \leq^{(*)} \sigma(k_{r-1}, p_a, p_0, n_{r-1}) \cdot \tau_1(k_r - k_{r-1}, p_a, p_0, n_r - n_{r-1}) =
\]
\[
 = \omega_r(k_r, k_{r-1}, p_a, p_0, n_r, n_{r-1}).
\]

Where inequality $(*)$ follows from Lemma~\ref{lemma:frac_sums_increasing}.
%the well-known property that if $a_i/b_i$ (for $a_i, b_i > 0$) is monotonicly increasing then
%(\sum_{i} a_i)/(\sum_{i} b_i)$ is also monotonicly increasing.

% 
% Now, if we consider $k^* = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ then it is easy to see that:
% 
% 
%  \[
%  k^{p_a, p_0, \alpha, k_{r-1}}_{min, [n_1, \ldots, n_r]}  
%   = \min\left\{k :
% \frac{
% s_{k_{r-1}, a, [n_{r-1}]}
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
% }{
% s_{k_{r-1}, 0, [n_{r-1}]}
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, 0, [n_r - n_{r-1}]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]
% \[
%   = \min\left\{k :
% \frac{
% s_{k_{r-1}, a, [n_{r-1}]}
% \left(
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
% + \sum_{i = k}^{n_r - n_{r-1}} s_{i, a, [n_r - n_{r-1}]}
% \right)
% }{
% s_{k_{r-1}, 0, [n_{r-1}]}
% \sum_{i = k}^{n_r - n_{r-1}} s_{i, 0, [n_r - n_{r-1}]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]

%  \[
%  k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} =  k^{p_a, p_0, \alpha}_{min, [n_j]}
%         = \min\left\{k :
% \frac{
% \sum_{i = k}^{n_j} s_{i, a, [n_j]}
% }{
% \sum_{i = k}^{n_j} s_{i, 0, [n_j]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]

% 
%  
%  Now, let us consider $j$-{th} round for $j = 1, \ldots, r$. We are going to show that
%  $k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} \leq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$.
%  
%  From Lemma~\ref{eq:kMin} we have:
%  \[
%  k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} =  k^{p_a, p_0, \alpha}_{min, [n_j]}
%   = \min\left\{k :
% \frac{
% \sum_{i = k}^{n_j} s_{i, a, [n_j]}
% }{
% \sum_{i = k}^{n_j} s_{i, 0, [n_j]}
% }\geq \frac{1}{\alpha}  \right\} =
% \]
% 
% From the definition of $k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ we have that for every $k \geq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$ it holds: $s_{k, a, [n_j]}/s_{k, 0, [n_j]} \geq 1/\alpha$ which means that:
% \[
% \frac{
% \sum_{i = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}}^{n_j} s_{i, a, [n_j]}
% }{
% \sum_{i = k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}}^{n_j} s_{i, 0, [n_j]}
% }\geq \frac{1}{\alpha}
% \]
% which means that $k^{p_a, p_0, \alpha}_{min, [n_1, \ldots, n_j]} \leq k^{p_a, p_0, \alpha}_{min, \BRAVO, n_r}$.

\end{proof}