### 2022-10-12 providence/sections/conc.tex

%Conclusions
A rigorous tabulation audit is an important part of a secure election. We present \Providence and demonstrate that it is as efficient as \Minerva and as flexible as \BRAVO. We present proofs and simulation results to verify the claimed properties of \Providence, and we provide an open source implementation of the stopping condition and useful related functionality for planning audits. We define the constraint of an acceptable probability of a misleading audit sample, and describe its importance to the planning process.