### 2023-02-18 providence/sections/intro.tex
Lots of changes

%Intro
It is well-known that electronic voting systems are vulnerable to software errors and manipulation which may be undetected. Errors and/or manipulation may not always change an election outcome, but we would want to know when they do. {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome. The {\em evidence-based elections} \cite{evidence-based} approach requires that election outcomes be supported by strong affirmative evidence. The evidence is generated during the election and publicly examined at the end, so that citizens may determine whether it provides strong support of the election outcome or is not sufficiently convincing. 
%by requiring not merely the use software independent systems to produce trustworthy evidence of outcome correctness; but also the examination of the evidence so that incorrect outcomes may be detected with high probability. 
One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely and procedures were followed; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.  
%Many US states have had pilot RLA programs. Additionally, some states allow RLAs to be used towards audit requirements, and some states require RLAs before elections can be certified. 
%We propose the \Providence audit, a new approach to a particular type of RLA---the ballot polling RLA described below---and propose a new model for the work load of an election. We show that \Providence is superior to the popular ballot polling RLA \Bravo for real elections, and describe the use of our open source implementation by the Rhode Island Board of Elections for an audit of their 2021 elections. \Providence was recently integrated as an option in Arlo \cite{arlo}, the most popular election audit software. %Our open-source implementations of \Providence and our audit planning tools are likely to be useful later this year; ballot polling audits are expected to be used as pre-certification RLAs in at least one statewide contest in both Georgia and Pennsylvania for the 2022 general elections in the US.

As a formalized approach to the examination of the evidence supporting vote tabulation, an RLA is an important part of an evidence-based election. When correctly used, it serves to greatly improve the trustworthiness of an election.

Significant efforts by nationwide organizations (such as Verified Voting, Brennan Center for Justice, Common Cause and Democracy Fund), local organizations, citizen advocates and experts have led to great progress towards the use of RLAs. Nonprofit VotingWorks has developed open-source election audit software, Arlo \cite{arlo}, and provides training in its use. Six states (Colorado, Georgia, Nevada, Pennsylvania, Rhode Island, Virginia) now require RLAs; three have statutory pilot programs (Indiana, Kentucky, Texas); four allow RLAs to satisfy a more general audit requirement (California, Ohio, Oregon, Washington); and two have an administrative pilot program (Michigan, New Jersey) \cite{vv_audits}. The effort to broaden the reach of RLAs continues, as election officials appear very keen to improve the trustworthiness of the elections they administer.

\subsection{Background on RLAs}
A tabulation audit will either end with a declaration that an election outcome is correct, or escalate to a full hand count. The integrity of an audit may be judged by how it deals with incorrect outcomes. A {\em risk-limiting audit (RLA)} guarantees a minimum probability that it will perform as it is supposed to (escalate to a full hand count) if the outcome is incorrect. Equivalently, this is also a guarantee of a maximum probability with which it will perform erroneously (declare the audit correct) when the outcome is incorrect. The {\em risk limit} of an RLA is the (guaranteed) maximum probability that an incorrect election outcome would be declared correct. Lower risk limits are better.

There are three types of RLAs: ballot comparison RLAs, batch comparison RLAs and ballot polling RLAs. This paper focuses on ballot polling RLAs which have been used in a number of US state pilots (California, Georgia, Indiana, Michigan, Ohio, Pennsylvania, Virginia and elsewhere), real statewide audits (Georgia, Virginia) \cite{vv_audits} and  audits of smaller jurisdictions, such as Montgomery County, Ohio \cite{usenix_minerva}. Ballot and batch comparison RLAs are described in section \ref{sec:related}.

\subsubsection{The Workflow of a Ballot Polling RLA}
\label{sec:polling}
%Ballot and batch comparison RLAs compare information gathered through manual examination of the sampled ballots with the corresponding information claimed by the voting system. For example, the information could be the vote on a particular ballot (stored by the voting system as a CVR or Cast Vote Record) or the tally of a particular box of ballots. 
A ballot polling RLA \cite{RLA} is based on the manual examination of the sampled ballots, and does not require any information from the tabulating system other than the tally. 
%, on the other hand, simply tallies the sampled ballots. It requires a much larger number of ballots than a ballot comparison RLA, but tends to be more feasible because it does not require any additional functionality of the voting system; in particular it does not require the system to generate or provide CVRs. 
More detail about the storage of ballots is required, however: a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example) to enable the sampling of specific ballots from the list.

All RLAs draw one or more ballots at a time; each such set of ballots is referred to as a {\em round}. We use notation and terminology from \cite{usenix_minerva,arxiv_athena,simulations,bravo} and also assume ballots are drawn with replacement.

Ballot polling audits proceed as follows. 
\begin{enumerate}
\item The ballot manifest is published. 
\item A first round size \cite{usenix_minerva} is chosen. %\fpo{the sentence is hard to understand}. pv: fixed
\item Ballots on the ballot manifest are sampled uniformly at random, with replacement, using a pseudorandom number generator, which is typically seeded by a natural source of randomness like rolling dice.
\item The physical ballots are found and manually interpreted, recording the manual interpretations. 
\item The stopping condition $\mathcal{A}$, a function of the manual interpretations of the current cumulative sample of ballots $X$, is computed. It outputs: 
\begin{enumerate}
\item \emph{Correct: stop the audit} or 
\item \emph{Undetermined: sample more ballots}. 
\end{enumerate}
Election officials may choose to abort this procedure and go to a full hand count at any time; we discuss this in more detail in section \ref{sec:related}.  
\item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
\end{enumerate}

Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required. For RLAs required by statute or legislation the successful completion of the RLA (or a full hand count confirming the outcome) is usually necessary before certification\footnote{If an RLA were performed after certification and determined that the outcome was incorrect, there might be no legal means of changing the outcome and this would significantly impact citizen confidence. For example, till recently, the state of Virginia required RLAs but they were to be performed after certification and could not be used to change an outcome. This was corrected through new legislation passed in April 2022.} and certification deadlines would play a large role in round size deternination.

%In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. This option is generally not incorporated \cite{bravo}. A manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary. 
% They would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence. 
%This value is the smallest risk limit of the RLA that would stop at the end of the round. 
%The stopping condition for \BRAVO, \Minerva and \Providence is described for a contest with two participants and formulated as a binary hypothesis test. Audits of multiple-candidate contests are treated as multiple audits of the pairwise contest between the announced winner and all other candidates. The \BRAVO and \Minerva stopping conditions are provided in section \ref{sec:model}.

%\BRAVO is also used for what are termed {\em stratified} audits [REF], where a single contest is audited using different approaches (ballot polling, ballot comparison, batch comparison) in different counties. When a ballot polling audit is used for the entire set of ballots, the null hypothesis is a tied outcome (the hardest incorrect outcome to detect). In a stratified audit, however, when testing individual units such as counties, one may need to assume a different null. This is because, for example, if a county with an announced large number of ballots for the winner were actually tied, the overall outcome would be incorrect. The proof of \Minerva's risk-limiting property assumes a tied null, and it is not immediately obvious how the proof may be modified to allow other possibilities. Thus, while \Minerva may be used for multi-candidate contests when all counties participate in the ballot polling audit, it cannot be used for stratified audits. \BRAVO, and \Providence, on the other hand, can.

%Using notation and terminology from , an audit $\mathcal{A}$ is a function that takes as input the current sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}. As mentioned earlier, we ignore a third option of a full hand count, leaving that decision to election officials, see section \ref{sec:related} for more detail.

\subsubsection{The Audit Model}
An audit is typically defined as a binary hypothesis test. If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}), we have the following definition of an RLA.

\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff for sample $X$
$$
Pr[\mathcal{A}(X) 
= \text{Correct} \,|\, H_0]\le \alpha
$$
\end{definition}

The definition above is valid at the end of the RLA, and not at the end of each round. 
%When $X$ is used to represent a sample while the RLA is ongoing, it is understood to be the cumulative sample, representing all the information obtained thus far about the underlying vote collection.

The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, have the audit declare an incorrect outcome as correct more often than guaranteed. An audit is an RLA only if there is a proof of its risk-limiting property. Thus, the adversary would need to, at the very least, invalidate an assumption of the proof to obtain:  
$$
Pr[\mathcal{A}(X) 
= \text{Correct} \,|\, H_0] > \alpha
$$

\subsection{Related Work}
\label{sec:related}
\input{sections/related}
%\subsection{Adversarial Model for RLAs}
%\label{sec:adv}
%Detailed descriptions of best practices for post-election audits may be found in \cite{best-practices,why-and-how}. For our purposes, we will assume that the best practices are followed: the paper trail consists of hand-marked paper ballots and is secured; a public compliance audit is carried out before the RLA to ensure that the processes for securing the paper trail were followed; voter authentication and registration processes are verified, and only legitimate voters cast no more than a single vote each; the risk-limiting audit is public. We will further assume that all software used in the RLA is open source and well-defined, so its output may be reproduced and thus verified by an observer wishing to do so with their own software.

%Referring to the ballot polling audit steps described in section \ref{sec:polling} above, we further assume that the ballot manifest is verified by the compliance audit; a secure PRNG is used; the seed is generated uniformly at random in a public process; the process of locating ballots is publicly observable and the located ballots can be viewed by the public. Because the PRNG is well-defined, as is the stopping condition, we may assume that the stopping condition is correctly computed, because it can be checked by the public through knowledge of the seed and the drawn ballots. Thus the only variable is round size.

%We define a {\em weak adversary} as one who can choose the first and subsequent round sizes before the audit begins and a {\em strong adversary} as one who can choose any round size at any time (before, of course, that particular round begins). In particular, a strong adversary may use information about the sample drawn thus far to decide the next round size, but a weak adversary may not.

\subsection{Gaps in Prior Work and Our Contributions}
Having provided context above, we now describe gaps in our knowledge of audits and describe our contributions. 
\subsubsection{Limitations of \Minerva}
Zag\'{o}rski {\em et al.}  prove that \Minerva \cite{usenix_minerva} is risk-limiting when the number of relevant ballots drawn in each round is pre-determined before any ballots are examined. They do not address the case of a stronger adversary (such as an audit insider) who can determine the size of the next round after knowing what votes are on the ballots sampled thus far. In particular, an open question about \Minerva is whether its RLA proof holds in this case. Can the audit insider increase the audit's error probability beyond its declared risk limit? Or is there no probabilistic adversarial advantage to being able to compute next round sizes after knowing the drawn sample? We do not answer this question, and to our knowledge, it remains open.

Until \Minerva is proven to be risk-limiting to a given risk limit for the adversary who can determine next round size after examining the current sample, it may not be used in audits whose round sizes are not pre-determined. This presents a major limitation, because the stopping probability of the next round is better estimated using information of the sample drawn thus far, but this would not be allowed for \Minerva. The current implementation of \Minerva integrated as an option in Arlo uses a fixed multiplier of the current round size to compute the next round size, thus allowing the first round to be computed as desired, and fixing the next round sizes thereafter\footnote{Note that every draw may contain invalid or irrelevant ballots, and thus the true number of relevant ballots can never be predetermined. However, because this is random, and not controllable by an adversary once the size of the draw is fixed, we assume that differences in the number of ballots average out, and that a fixed draw size is sufficient, though this is not explicitly proven in \cite{usenix_minerva}.}. This could lead to the drawing of too few or too many ballots and greatly constrains audit planning.

% In MINERVA, the number of ballots drawn in each round is determined before any ballots are drawn. Because invalid ballots and ballots that are inconsequential for the contest being audited would be drawn in addition to relevant ballots, the assumption used by the proof is not true in general. (We are grateful to Philip Stark for drawing our attention to this.) However, any variation in number of relevant ballots drawn for a fixed round size would be random and not chosen by an adversary; the proof showing the risk-limiting property of MINERVA could hence be extended.

The risk limit for B2, EoR and SO \BRAVO is fixed independent of whether next round sizes are determined with or without knowledge of the current sample. This allows \BRAVO audits the flexibility of choosing smaller subsequent round sizes if the sample drawn so far is a ``good'' sample. An open question is whether a ballot polling RLA exists with the efficiency of \Minerva and this flexibility of \BRAVO.

\subsubsection{Limitations in Existing Workload Measures}
A major limitation of our understanding of the ballot polling problem as a community is that we use the number of ballots drawn or values proportional to this number \cite{mclaughlin_thesis,bernhard-diss,RI-report} as measures of the workload of an audit. If this were a correct measure of the workload of an audit, we would want to use B2 audits (round size is one) and make decisions about stopping the audit after drawing each ballot, because this leads to the smallest expected number of ballots. As described above, election officials, on the other hand, greatly prefer drawing many ballots at once. From conversations with election officials and staff members of Verified Voting, Brennan Center and Common Cause who have been training election officials to perform RLAs, we estimate that this preference is likely due to the following.

Firstly, each round has an overhead workload as well, including setting up the round and communicating among the various localities involved in conducting the audit (for example, audits of statewide contests involve the drawing of ballots at county offices where the ballots are stored). Secondly, there is an overhead to finding a storage box and unsealing it. For large round sizes, multiple ballots may be drawn at once from a box, and the number of boxes retrieved is smaller than the number of ballots (storage boxes commonly contain many hundreds of ballots each). 
%For smaller round sizes, the number of times a box is retrieved would be roughly identical to the number of ballots drawn, as it is unlikely that a single box will hold multiple ballots from the sample. 
Finally, in the current environment of misinformation, election officials fear a misleading audit sample (falsely indicating that the loser won), preferring to structure audits to reduce the chances of such samples, thus implicitly choosing larger round sizes.

Thus the workload of an audit is not simply a linear (or affine) function of the number of ballots drawn. Relatedly, an optimal round schedule is not completely determined by the expected number of ballots drawn. It depends on other variables as well. The consideration of all these variables is necessary while planning an audit.

\subsubsection{Our Contributions}
Our primary contribution is a new RLA, \Providence, which gives the efficiency of \Minerva and is also resistant to an adversary who can choose the next round size after knowing the current sample. 
%Additionally, \Providence does not assume that the null hypothesis is a tied election and hence is more broadly useful than \Minerva. 
\BRAVO is a very flexible audit enabling audit planning; with the introduction of \Providence, similar planning is possible with greater efficiency. In a contest with a narrow margin (in the 2020 US Presidential election, eight states had margins smaller than 0.03) the difference in number of ballots sampled using \Providence over \BRAVO could correspond to many days of work which would need to be completed before a certification deadline.

The stopping condition for \Minerva does not take into account the sample obtained in previous rounds, and, in \cite{usenix_minerva}, its risk limit is estimated through weighted averages across multiple rounds, assuming that round sizes do not depend on the previous sample. 
%Further, it assumes that the null hypothesis is a tie. 
A far simpler proof is possible for the risk-limiting property of \Providence. 
In particular, the proof does not require an assumption about round sizes.

We provide the following:
\begin{enumerate}
\item Proof that \Providence is an RLA and resistant to an adversary who can choose next round sizes after knowing the current sample.
\item Simulations of \Providence, \Minerva, SO \BRAVO, and EoR \BRAVO which show that \Providence uses number of ballots similar to those of \Minerva, both fewer than either version of \BRAVO.
\item Results and analysis from the use of \Providence in a pilot audit in Rhode Island.
\item A model of workload that includes the overhead effort of each round and the overhead effort of retrieving a storage unit of ballots; simulations that illustrate the use of this model to compare the different types of ballot polling audits and to plan an audit with minimal workload.
\item An analysis of round size as a function of the maximum acceptable probability of a misleading audit sample.
\item Open source implementation of \Providence and audit planning tools. 
%including the novel metric Probability of Misleading(name?)
\end{enumerate}

\Providence may be used in any audit where sampling is with replacement and audited contests may be expressed as pairwise plurality contests: for example, it can be used in plurality and majoritarian elections.

%Our results demonstrate the superiority of \Providence over the other audits. Our work may be used by election officials to plan ballot polling audits, including in Georgia and Pennsylvania in 2022.

\subsubsection{Organization} 
Sections \ref{sec:model} provides preliminaries on the \BRAVO and \Minerva audits. Section \ref{sec:prov} describes the \Providence audit, section \ref{sec:sims} the simulations comparing the number of ballots drawn using various ballot polling audits and section \ref{sec:pilot} the use of \Providence in an audit carried out by the Board of Elections of Rhode Island. Section \ref{sec:workload} presents our workload model and describes its use for a ballot polling audit using details of the 2020 US Presidential election in the state of Virginia. Our conclusions, the availability of an audit implementation and acknowledgements may be found in sections \ref{sec:conc}, \ref{sec:avail} and \ref{sec:ack} respectively.