### 2023-02-19 providence/Extra.tex
Example ongoing

We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit\footnote{Technically, it is the maximum risk, corresponding to the election outcome hardest to detect.}. Loosely speaking, it is the sum of the probabilities of every sequence that stops the audit. The sequences that stop the audit depend on the audit, and not on the underlying vote distribution as the audit does not know what it is. The probabilities used for risk contributions are computed assuming the null hypothesis; those used for stopping probability computations assume the announced tally. Thus, every time an audit stops, it incurs a contribution to its risk and its stopping probability. Audit design determines, among other things, how the risk and stopping probability contributions are distributed across rounds, as long as the overall risk is limited by $\alpha$.